\sysfull

### 2018-08-05 paper/asplos19/evaluation.tex
partial commit

% \begin{wrapfigure}{t!}{0.5\textwidth}
\begin{figure}
	%\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
    \caption{Breakdown of overhead for the proposed coalescing strategies
relative to the baseline without coalescing.
NC: no coalescing, EO: energy-oblivious coalescing, HG: history-guided coalescing, WHG:
energy- and task-aware coalescing. All coalescing strategies reduce total
overhead and maximize useful work. HG and WHG perform better than EO.}
	\label{fig:overallOverheadBreakdown}
\end{figure}
% \end{wrapfigure}

Our evaluation quantitavely demonstrates that \sys (i) \emph{reduces the memory
protection overhead} by coalescing, (ii) \emph{improves execution speed} for
conservative static task decompostions compared to state-of-the-art runtimes; %
and (iii) \emph{provides computation progress} at the sub-task granularity,
preventing non-termination at adverse energy conditions where state-of-the-art
runtimes fail. We measured the execution time of the benchmarks implemented
using \sys with each variant of the two proposed adaptation mechanisms:
\emph{task coalescing} and \emph{task downsizing}. We compared \sys's best
configurations against a state-of-the art system, Alpaca~\cite{alpaca}. We have
also broken down \sys's overhead to explain the source of \sys's improved
performance.

\subsection{Characterization of Overhead}
\label{sec:coala_overhead}

%\indent \textbf{Overhead Reduced by Coalescing.}

@@

For each coalescing strategy from Section~\ref{sec:task_adaptation} (EO, HG,
and WHG) and for a baseline without coalescing (NC), we have measured the time
spent on (i) executing useful task code, including \sys's overhead on access to
protected variables, (ii) re-executing task code after a power failure, and
(iii) committing state to non-volatile memory at the end of each (coalesced)
task.  In this experiment the WISP was 15\,cm away from the signal generator
antenna.
%
The page size was set to the best performing value (32~B for \textbf{bc} and
256~B pages for the rest) identified in detailed evaluation of paging in
Section~\ref{sec:results_memory_management}.

The overhead incurred by each coalescing strategy is broken down in
Figure~\ref{fig:overallOverheadBreakdown}. Without coalescing enabled (NC), the
re-execution penalty is smallest, because the amount of work that can happen
between commits and may have to be re-executed if interrupted is smallest when
work from multiple static tasks is not combined.
%
However, any gain from a reduced re-execution penalty is cancelled out by
the increased commit overhead that is incurred at the end of each static
task.
%
Across all benchmarks, all coalescing strategies reduce more commit overhead
than the re-execution overhead they add.
%
This net overhead reduction is greatest in HG and WHG strategies compared
to the EO strategy. We attribute this discrepancy to EO's slow adjustment
of the target task size without regard to execution history.
%
In the subsequent experiments, we focus on the better-performing HG and WHG
strategies.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/overallOverhead}
	\includegraphics[width=\columnwidth]{figures/memAccess}
	\caption{\sys kernel overhead: a breakdown of time spent in
the application and in the runtime (left) and a breakdown of protected memory
accesses into three types with different costs. NC: no coalescing, HG:
history-guided coalescing, WHG: weighted history-guided coalescing.}
	\label{fig:coalEfficiency}
\end{figure}

%\textbf{\sys Kernel Overhead Breakdown.}

\subsubsection{\sys Kernel Overhead Breakdown}

Figure~\ref{fig:coalEfficiency} (left) breaks down the time spent on executing
useful task code and committing further into time spent by the application and
time spent by \sys's runtime.
%
This dataset is generated in a separate experiment with the same setup as the
previous experiment in Section~\ref{sec:overhead-coalescing}.
%
Consistent with the previous results in
Figure~\ref{fig:overallOverheadBreakdown}, commit overhead is highest when
coalescing is not used to reduce the number of commits.
%
However, coalescing increases the cost of each access to a protected variable,
since work in the address translation in the paging system increases as the
task size grows.
%
For both coalescing strategies (HG and WHG) accesses to protected variables
constitute about 30\% of the runtime overhead. Dynamic address translation
necessary on each protected access is the most critical bottleneck for \sys.

%\textbf{Protected Memory Accesses Breakdown.}

\subsubsection{Protected Memory Accesses Breakdown}

Figure~\ref{fig:coalEfficiency} (right) breaks down protected memory accesses
into three categories.
%
Each type of protected access incurs a different amount of overhead: access to
the same page in SRAM as the previous access is cheapest, access to a different
SRAM page has a larger cost, and access to a page that needs to be swapped in
from FRAM into SRAM is most expensive.
%
The results in the figure show that the overwhelming majority of accesses are
of the cheapest kind, which motivated us to optimize this access type in our
implementation (cf. Section~\ref{sec:impl:paging}).
%
The search for the accessed page in the \texttt{working} buffer,
will return immediately when the page is the most recently accessed page.
%
Only \textbf{cuckoo}, \textbf{dijkstra}, and \textbf{fft} have non-negligible
number of accesses to a different SRAM page, which is due to the larger working
set and a less regular access pattern in these applications.
%
In general, memory access patterns are shaped by the application, and the more
program state is protected, the higher the rate of page swaps.

\subsection{Execution time}
\label{sec:result_coalescing}

Having shown in Section~\ref{sec:coala_overhead} that coalescing reduces
overhead, we now investigate the outcome of this reduction on the total
execution time. We first investigate difference variants of \sys and
then compare the best variant to Alpaca.

%\textbf{\sys Speedup with Coalescing.}

\subsubsection{\sys Speedup with Coalescing}

Figure~\ref{fig:coalescing} shows the \sys run time with two coalescing
strategies (HG, WHG) normalized to the run time without coalescing (NC).
%
The experimental setup is the same as that in Section~\ref{sec:coala_overhead}.
%
The results in Figure~\ref{fig:coalescing} (left) show that all benchmarks
complete faster with coalescing than without coalescing: from 25\% faster
(\textbf{ar}) up to 70\% faster (\textbf{sort}).
%
This speedup is a consequence of the reduced overhead demonstrated in
Section~\ref{sec:coala_overhead}.
%
However, the magnitude of the speedup is (1) highly application-dependent and
(2) largely similar across the two coalescing strategies, with the exception of
\textbf{fft}.
%
In some cases (\textbf{bc}, \textbf{cuckoo}, \textbf{sort}) the history-guided
feature in the coalescing algorithm is counter-productive.
%
This occurs in task decompositions with a uniform task energy, where counting
tasks disregarding their energy provides an equal amount of information with a
smaller effort.
%
In \textbf{fft}, tasks are not uniform, and accounting for task energies is
beneficial. In fact, the lack of task energy awareness is detrimental: with HG
\textbf{fft} runs slower than without any coalescing (NC).
%
The speedup is highest for \textbf{bc}, \textbf{cuckoo}, \textbf{dijkstra} and
\textbf{sort}, because their tasks are small and are easily coalesced,
eliminating many unnecessary commits.

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/coalStrategies}%
    \caption{\sys's coalescing performance. Left: Application execution time
with coalescing strategies (HG and WHG) normalized to the execution time
without coalescing.}
	\label{fig:coalescing}
\end{figure}

@@

\subsubsection{Benefits of Adaptive Tasks}

We now compare \sys's performance to Alpaca~\cite{alpaca}---a
\emph{non-adaptive} task-based system with tasks fixed at compile time.
Figure~\ref{fig:runtime} (right) shows the average execution time of each
application for \sys and Alpaca, normalized to the latter. The results show
that \sys provides a performance benefit compared to Alpaca for most
applications, with the exception of \textbf{ar} and \textbf{cuckoo} at some
distances.  The speedup is greatest for applications with repeated write after
read dependencies throughout their code, particularly involving arrays
(\textbf{dijkstra}, \textbf{fft} and \textbf{sort}). \sys's virtualized memory
successfully amortizes the overhead of protecting memory that is accessed in
such patterns.  In applications without locality among accesses to protected
variables, including array elements, \sys incurs overhead from memory
virtualization that causes its performance to be comparable to (or worse than)
Alpaca (\textbf{ar}, \textbf{cuckoo}).

@@

\begin{figure}
	\centering
	\includegraphics[width=0.49\columnwidth]{figures/page_exec-time}
	\includegraphics[width=0.49\columnwidth]{figures/pagePulls}
    \caption{Effect of page size on execution time (left) and on the number of
page faults (right). The execution time is normalized to the lowest in each set
of page sizes.}
	\label{fig:page_size}
\end{figure}

>>>>

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
    \caption{Breakdown of overhead for the proposed coalescing strategies
relative to the baseline without coalescing.
NC: no coalescing, EO: energy-oblivious coalescing, HG: history-guided coalescing, WHG:
energy- and task-aware coalescing. All coalescing strategies reduce total
overhead and maximize useful work. HG and WHG perform better than EO.}
	\label{fig:overallOverheadBreakdown}
\end{figure}
% \end{wrapfigure}
%
Building on top of the existing work, we have selected the most efficient run-time library~\cite{alpaca} and compared the performance of \sys against it.
Our evaluation quantitatively demonstrates that \sys (i) \emph{reduces the memory
protection overhead} by coalescing, (ii) \emph{improves execution speed}, in most cases; %
and (iii) by mean of task-downscaling \sys is able to progress where static system stuck in task non-termination loop. 
We have also broken down \sys's overhead to explain the source of \sys's improved performance.
%
\subsection{Characterization of Overhead}
\label{sec:coala_overhead}

@@

For each coalescing strategy from Section~\ref{sec:task_adaptation} (EO, EG,
and WEG) and for a baseline without coalescing (NC), we have measured the time
spent on (i) executing useful task code, including \sys's overhead on access to
protected variables, (ii) re-executing task code after a power failure, and
(iii) committing state to non-volatile memory at the end of each (coalesced)
task.  In this experiment the WISP was 15\,cm away from the signal generator
antenna.
%
The page size was set to the best performing value (32~B for \textbf{bc} and
256~B pages for the rest) identified in detailed evaluation of paging in
Section~\ref{sec:results_memory_management}.

The overhead incurred by each coalescing strategy is broken down in
Figure~\ref{fig:overallOverheadBreakdown}. Without coalescing enabled (NC), the
re-execution penalty is smallest, because the amount of work that can happen
between commits and may have to be re-executed if interrupted is smallest when
work from multiple static tasks is not combined.
%
However, any gain from a reduced re-execution penalty is canceled out by
the increased commit overhead that is incurred at the end of each static
task.
%
Across all benchmarks, all coalescing strategies reduce more commit overhead
than the re-execution overhead they add.
%
This net overhead reduction is greatest in EG and WEG strategies compared
to the EO strategy. We attribute this discrepancy to EO's slow adjustment
of the target task size without regard to the energy conditions.
%
In the subsequent experiments, we focus on the better-performing EG and WEG
strategies.
%
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/overallOverhead}
	\includegraphics[width=\columnwidth]{figures/memAccess}
	\caption{\sys kernel overhead: a breakdown of time spent in
the application and in the runtime (top) and a breakdown of protected memory
accesses into three types with different costs. NC: no coalescing, HG:
history-guided coalescing, WHG: weighted history-guided coalescing.}
	\label{fig:coalEfficiency}
\end{figure}
%
\subsubsection{\sys Kernel Overhead Breakdown}
%
Figure~\ref{fig:coalEfficiency} (top) breaks down the time spent on executing
useful task code and committing further into time spent by the application and
time spent by \sys's runtime.
%
This dataset is generated in a separate experiment with the same setup as the
previous experiment in Section~\ref{sec:overhead-coalescing}.
%
Consistent with the previous results in
Figure~\ref{fig:overallOverheadBreakdown}, commit overhead is highest when
coalescing is not used to reduce the number of commits.
%
However, coalescing increases the cost of each access to a protected variable,
since work in the address translation in the paging system increases as the
task size grows.
%
For both coalescing strategies (EG and WEG) accesses to protected variables
constitute about 30\% of the runtime overhead. Dynamic address translation
necessary on each protected access is the most critical bottleneck for \sys.
%
\subsubsection{Protected Memory Accesses Breakdown}
%
Figure~\ref{fig:coalEfficiency} (bottom) breaks down protected memory accesses
into three categories.
%
Each type of protected access incurs a different amount of overhead: access to
the same page in SRAM as the previous access is cheapest, access to a different
SRAM page has a larger cost, and access to a page that needs to be swapped in
from FRAM into SRAM is most expensive.
%
The results in the figure show that the overwhelming majority of accesses are
of the cheapest kind, which motivated us to optimize this access type in our
implementation (cf. Section~\ref{sec:impl:paging}).
%
Only \textbf{cuckoo}, \textbf{dijkstra}, and \textbf{fft} have non-negligible
number of accesses to a different SRAM page, which is due to the larger working
set and a less regular access pattern in these applications.
%
In general, memory access patterns are shaped by the application, and the more
program state is protected, the higher the rate of page swaps.
%
\subsection{Execution time}
\label{sec:result_coalescing}
%
Having shown in Section~\ref{sec:coala_overhead} that coalescing reduces
overhead, we now investigate the outcome of this reduction on the total
execution time. We first investigate difference variants of \sys and
then compare the best variant to Alpaca~\cite{alpaca}.
%
\subsubsection{\sys Speedup with Coalescing}

Figure~\ref{fig:coalescing} shows the \sys run time with two coalescing
strategies (EG, WEG) normalized to the run time without coalescing (NC).
%
The experimental setup is the same as that in Section~\ref{sec:coala_overhead}.
%
The results in Figure~\ref{fig:coalescing} show that all benchmarks
complete faster with coalescing than without coalescing: from 25\% faster
(\textbf{ar}) up to 70\% faster (\textbf{sort}).
%
This speedup is a consequence of the reduced overhead demonstrated in
Section~\ref{sec:coala_overhead}.
%
However, the magnitude of the speedup is (1) highly application-dependent and
(2) largely similar across the two coalescing strategies, with the exception of
\textbf{fft}.
%
In some cases (\textbf{bc}, \textbf{cuckoo}, \textbf{sort}) the Weight Energy-guided
feature in the coalescing algorithm is counter-productive.
%
This occurs in task decompositions with a uniform task energy, where counting
tasks disregarding their energy provides an equal amount of information with a
smaller effort.
%
In \textbf{fft}, tasks are not uniform, and accounting for task energies is
beneficial. In fact, the lack of task energy awareness is detrimental: with EG
\textbf{fft} runs slower than without any coalescing (NC).
%
The speedup is highest for \textbf{bc}, \textbf{cuckoo}, \textbf{dijkstra} and
\textbf{sort}, because their tasks are small and are easily coalesced,
eliminating many unnecessary commits.
%
\begin{figure}
	\includegraphics[width=\columnwidth]{figures/coalStrategies}%
    \caption{\sys's coalescing performance.  Application execution time
with coalescing strategies (EG and WEG) normalized to the execution time
without coalescing.}
	\label{fig:coalescing}
\end{figure}

@@

\subsubsection{Benefits of Adaptive Tasks}
%
We now compare \sys's performance to Alpaca~\cite{alpaca}---a
\emph{non-adaptive} task-based system with tasks fixed at compile time.
Figure~\ref{fig:runtime} (right) shows the average execution time of each
application for \sys and Alpaca, normalized to the latter. The results show
that \sys provides a performance benefit compared to Alpaca for most
applications, with the exception of \textbf{ar} and \textbf{cuckoo} at some
distances.  The speedup is greatest for applications with repeated write after
read dependencies throughout their code, particularly involving arrays
(\textbf{dijkstra}, \textbf{fft} and \textbf{sort}). \sys's virtualized memory
successfully amortizes the overhead of protecting memory that is accessed in
such patterns.  In applications without locality among accesses to protected
variables, including array elements, \sys incurs overhead from memory
virtualization that causes its performance to be comparable to (or worse than)
Alpaca (\textbf{ar}, \textbf{cuckoo}).

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/page_exec-time}
	\includegraphics[width=\columnwidth]{figures/pagePulls}
    \caption{Effect of page size on execution time (left) and on the number of
page faults (right). The execution time is normalized to the lowest in each set
of page sizes.}
	\label{fig:page_size}
\end{figure}
